### 2017-03-06 paper/paper.Rmd
more on methods

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: '~/Google Drive/INFO498C/INFO498Final/paper/svm-latex-ms.tex'
title: "Assesing Health Care Coverage and Access Utilizing the National Health Interview Survey"
author:
- name: John Diego
  affiliation: University of Washington
- name: Warren Wakuzawa
  affiliation: University of Washington
- name:  Adrian Santiago
  affiliation: University of Washington
abstract: "The National Health Interview Survey (NHIS) is the nation’s largest in-person household health survey. It has been conducted annually since 1957 by the National Center for Health Statistics (NCHS), which is a part of the Centers for Disease Control and Prevention (CDC). A broad range of topics are covered but we will specifically focus on healthcare access and expenditure across a multitude of factors such as employer information, race, and academic background."
keywords: "healthcare, nhis, aca, medicaid, medicare"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 10pt
# spacing: double
bibliography: ~/Google Drive/INFO498C/INFO498Final/paper/master.bib
biblio-style: apsr
---

# INTRODUCTION
The Patient Protection and Affordable Care Act (PPACA), commonly shortened to the Affordable Care Act (ACA) and nicknamed Obamacare, is one of the most important healthcare legislature, creating a significant impact on the US Health care system. However, with the arrival of President Trump in the Oval Office, and with the ACA repeal now underway, some speculate huge consequences such as increased death rates([@trump]). In light of these recent events, it is important that we take a look at the current state of our healthcare system and question whether it is failing or not.

The purpose of this paper is to assess healthcare coverage and access across a  multitude of factors. We will specifically look at three aspects of health care - the first aspect is the relationship between educational attainment and access to medical insurance, the second is an overall look at the population on which had health care coverage, and the final exploration covers healthcare expedinture over the years.

# RELATED WORK
A description of previous papers or projects related to your project.

# METHODOLOGY
To answer the questions we layed out, we use data provided from the NHIS survey. We used various subsets of the survey data, depending on the specific question being analyzed.

## **Educational Attainment Vs Healthcare Access**

## **Health Care Access**

## **Health Care Expedentiture**
To gain a general outlook on the healthcare system, one way of looking at it is through the perspective of expenditures. Specifically the question we wanted to explore is, _"How has the amount of money spent for medical care changed over the years?"_ One specific variables that was counted for from the survey was **"Amount family spent for medical care, past 12 months"** (N = ). This variable only provided a rough estimate on the family's expenditure on medical care, but is still relevant since interviewers directed respondents to exclude the insurance premiums, over-the-counter drugs, and any costs for which they expected to be reimbursed. Using the provided calculated weight and strata, a **stratified random sampling** was implemented to calculate a weighted score.

# RESULTS
The major insights resulting from your research.

## **Plots**

# DISCUSSION
What are the implications of your results?

# FUTURE WORK
A description of how your system could be extended.