### 2017-03-08 paper/index.Rmd
update app and incude html of graphs

---
title: "healthcare access"
author: "John Diego, Warren Wakuzawa, Adrian Santiago"
date: "2/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 9.5)
source("calculations.R")

library(ggplot2)
```

## Overview of Healthcare Access and Coverage in the US

```{r expenddistribution, echo=FALSE}
# Line chart
# Change of expend over years (Not weighted)
ggplot(spend.years, aes(x=YEAR, y=count/1000, colour=HCSPENDY)) + geom_line() +
  geom_point(size=.7) +
  labs(x="Year", y="Sum of People (Thousands)", title = "Distribution of Health Expenditure") +
  theme(legend.title=element_blank())
```

```{r weightedexpendproportion, echo=FALSE, fig.height=6}
ggplot(hi.spend, aes(y=factor(YEAR))) +
  geom_point(data= hi.spend, aes(x=zero/population * 100, colour='red'), size=.9) +
  geom_errorbarh(data= hi.spend, aes(x=zero/population * 100, xmin=ci_u.zero/population * 100, 
                                     xmax=ci_l.zero/population * 100),height=0.1) +
  geom_point(data= hi.spend, aes(x=lessthan500/population * 100, colour='gold'), size=.9) +
  geom_errorbarh(data= hi.spend, aes(x=lessthan500/population * 100, xmin=ci_l.lessthan500/population * 100, 
                                     xmax=ci_u.lessthan500/population * 100),height=0.1) +
  geom_point(data= hi.spend, aes(x=spend3000to4999/population * 100, colour='blue'), size=.9) +
  geom_errorbarh(data= hi.spend, aes(x=spend3000to4999/population * 100, xmin=ci_l.spend3000to4999/population * 100, 
                                     xmax=ci_u.spend3000to4999/population * 100),height=0.1) +
  coord_flip() + 
  scale_color_discrete(name="Expenditure", labels = c("$3000 to $4999", "Less than $500", "Zero")) +
  labs(x="Weighted Proportion of People (%)", y="Year", title="Weighted Sum Proportion of Health Care Expenditure")
```

```{r healthcovereduattainment, echo=FALSE}
ggplot(d2015, aes(EDUC1)) + geom_histogram(stat='count') + facet_wrap(~NOTCOV) + 
  ggtitle('Education Level vs. Health Coverage') + labs(x='Education Level') 
```

```{r workhealthinsuranceeduattainment, echo=FALSE}
ggplot(d2015, aes(EDUC1)) + geom_histogram(stat='count') + facet_wrap(~HIEMPOF) +
  ggtitle('Education Level vs. Workplace-offered Health Insurance') + labs(x='Education Level')
```